---
title   : BMR Calculator 
subtitle: Developing Data Products Course Project
author  : Daniela Petruzalek
job     : Data Scientist
widgets : mathjax
mode    : selfcontained
---

## Introduction

Basal Metabolic Rate (BMR) is the minimal rate of energy expenditure per unit time by endothermic animals at rest. It is an important metric for fitness enthusiasts and people seeking life changing dietary goals.

The BMR added to the active calories (calories expended during daily activities and exercises) accounts for the total calorie consumption of the body during a 24 hour period. The classical approach to weight-loss is to ingest less calories from the diet than the total calorie consumption during the day.

The food labels generally account for a daily intake of 2000 calories, but this number is not ideal for every person. This project, the BMR Calculator, helps the user to calculate it's own BMR using three different models. The estimated BMR value is multiplied by a number that corresponds to the individual's activity level. The resulting number is the recommended daily kilocalorie intake to maintain current body weight.

---

## The Original Harris-Benedict Equation

The Harris–Benedict equation (also called the Harris-Benedict principle) is a method used to estimate an individual's basal metabolic rate (BMR) and daily kilocalorie requirements. The equations for each gender are given below.

1. For men: $$P = (\frac{13.7516 m}{1 kg}+\frac{5.0033 h}{1 cm}-\frac{6.7550 a}{1 year}+66.4730)\frac{kcal}{day}$$

2. For women: $$P = (\frac{9.5634 m}{1 kg}+\frac{1.8496 h}{1 cm}-\frac{4.6756 a}{1 year}+655.0955)\frac{kcal}{day}$$

```{r, echo=FALSE}
weight <- 65
height <- 163
age    <- 31
P <- 9.5634 * weight + 1.8496 * height - 4.6756 * age + 655.0955
```

__Example:__ For a women with `r age` years old, `r height` cm and `r weight` kg, the calculated BMR is: `r round(P,2)`. If she has an activity level of light exercise (1-3 days per week), her total daily calorie consumption is `r round(P * 1.375, 2)` kcal.

---

## The Revised Harris-Benedict Equation

In 1984, the original Harris-Benedict equations were revised using new data. In comparisons with actual expenditure, the revised equations were found to be more accurate.

1. For men: $$P = (\frac{13.397 m}{1 kg}+\frac{4.799 h}{1 cm}-\frac{5.677 a}{1 year}+88.362)\frac{kcal}{day}$$

2. For women: $$P = (\frac{9.247 m}{1 kg}+\frac{3.098 h}{1 cm}-\frac{4.330 a}{1 year}+447.593)\frac{kcal}{day}$$

```{r, echo=FALSE}
weight <- 65
height <- 163
age    <- 31
P <- 9.247 * weight + 3.098 * height - 4.330 * age + 447.593
```

__Example:__ For a women with `r age` years old, `r height` cm and `r weight` kg, the calculated BMR is: `r round(P,2)`. If she has an activity level of light exercise (1-3 days per week), her total daily calorie consumption is `r round(P * 1.375, 2)` kcal.

---

## The Mifflin St Jeor Equation

This model was introduced in 1990 and is considered most accurate than the Revised Harris-Benedict equations.

1. For men: $$P = (\frac{10.0 m}{1 kg}+\frac{6.25 h}{1 cm}-\frac{5.0 a}{1 year}+5)\frac{kcal}{day}$$

2. For women: $$P = (\frac{10.0 m}{1 kg}+\frac{6.25 h}{1 cm}-\frac{5.0 a}{1 year}-161)\frac{kcal}{day}$$

```{r, echo=FALSE}
weight <- 65
height <- 163
age    <- 31
P <- 10.0 * weight + 6.25 * height - 5.0 * age - 161
```

__Example:__ For a women with `r age` years old, `r height` cm and `r weight` kg, the calculated BMR is: `r round(P,2)`. If she has an activity level of light exercise (1-3 days per week), her total daily calorie consumption is `r round(P * 1.375, 2)` kcal.
